---
title: "Italian Cluster Analysis - Cleaning"
author: "Gina Cuomo-Dannenburg, Katy Gaythorpe and Arran Hamlet"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
    fig_width: 7
    fig_height: 5
    fig_caption: true
    highlight: "tango"
    reference_docx: word_style.docx
---

This is the data on the a subset of 3005 individuals from a village in Italy. It includes their personal details, along with some information on symptoms (which isn't really relevant for this analysis), and information on contact history. 

The villagers were tested at two time points -- I have inferred (from what I can understand without speaking Italian). The first test occurred on February 21st and the second on March 3rd. We are only interested in actual cases so I will filter these by those who test positive at any point. I'm presuming that the additional date columns is the date of confirmation rather than anything else.

```{r read-data-file}
set.seed((1))

## fix encoding of characters
full_data <- readxl::read_excel("final.data.xlsx", sheet = 1,
                                               guess_max = 2900)

head(full_data)
unique(full_data$"43882")
```

```{r contact-data-set}
## find a way of doing this that doesn't require the specification of columns --> patient names
contact_data_long <- tidyr::gather(full_data, contact_id, contact_type, 
                     "UQfjwKik":"db2cfcea_place", factor_key = TRUE) %>%
  dplyr::filter(!is.na(contact_type) == TRUE) 

contact_data_long <- contact_data_long[,c("id", "household_id", "contact_id", "contact_type")]
contact_ids <- unique(contact_data_long$contact_id)

missed_ids <- setdiff(contact_ids, full_data$id)

### check none of these are in fact individual entries in linelist
if(sum(missed_ids %in% full_data$id) != 0) {
  stop("Some of the additional ids are in the full data already")
}

missing_households <- data.frame(id = missed_ids, 
                                 household_id = c(rep("out_of_vo", length(missed_ids) - 3), 
                                                  rep("location_in_vo", 3)))
household_df <- full_data[,c("id", "household_id")]
household_complete <- rbind(household_df, missing_households)
names(household_complete) <- c("contact_id", "contact_household")

contact_data <- dplyr::full_join(contact_data_long, household_complete, by = "contact_id") %>%
  dplyr::filter(contact_type != "NA")


```

```{r adding-household-contacts}
length(unique(household_df$household_id))

## if two individuals live togeher, add them as contacts
household_contacts <- household_df %>%
  dplyr::group_by(household_id) %>%
  dplyr::mutate(individuals = list(id)) %>%
  dplyr::filter(length(individuals) > 1) 

## testing getting out permutations
combn(c("e5249500", "ca2448ad", "2ebf8b2e", "c5c409c4"), m = 2)
combn(c("e5249500", "ca2448ad"), m = 2)
class(household_contacts$individuals)

permute <- function(x) {
  combn(x, m = 2)
  }

test <- c(lapply(household_contacts$individuals, FUN = permute))
head(test)

test <- as.data.frame(test)

household_contacts_long <- t(test)
household_contacts %<>% as.data.frame()
household_contacts_long %<>% as.data.frame()

names(household_contacts_long)[1] <- "id"

test <- dplyr::left_join(household_contacts_long, household_contacts, by = "id")
names(test)[c(1:2, 4)] <- c("to", "from", "to_household")

test %<>% dplyr::mutate(from_household = to_household) %>%
  dplyr::mutate(contact_type = 4)

## keep the columns we want
household_data_final <- test[,c("to", "from", "household_id", "contact_type")]

```

```{r cleaning-household-contacts}
## check that this matches the rest of the contact dataset
head(contact_data)
names(household_data_final) <- c("id", "contact_id", "household_id", "contact_type")
household_data_final$contact_household <- household_data_final$household_id

contact_data <- rbind(contact_data, household_data_final)

# remove duplicates (to --> from = from --> to) -- we are not interested in the type of contact
df_sort <- t(apply(contact_data[,1:2], 1, sort))
contact_final <- contact_data[!duplicated(df_sort),]

## checking that duplicates have been removed
if(nrow(contact_data) == nrow(contact_final)) {
  stop("No duplicates removed")
}

```


```{r write-contacts-csv}
## contact type 3 = indirect contacts
direct_contact <- contact_final %>% dplyr::filter(contact_type != 3)

write.table(contact_final,"all_contacts.tsv", quote=FALSE, sep='\t', row.names = FALSE)
write.table(direct_contact,"direct_contacts.tsv", quote=FALSE, sep='\t', row.names = FALSE)
```

```{r gather-test-dates}
full_data$id <- as.factor(full_data$id)

## filter for only cases who tested positive
data <- dplyr::filter(full_data, positive == TRUE)

## starts with 4 == the date of testing (this is an artefact of reading in an excel file)
long_data <- tidyr::pivot_longer(data = data, cols = starts_with("4"), names_to = "date_confirmed",
                                 values_to = "test_result") %>%
  ## another artefact of reading in from xlsx -- usually would use !is.na() 
  dplyr::filter(test_result != "NA")

## break clause to check we keep all the entries
if (nrow(data) != length(unique(long_data$id))) {
  stop("Entries have been lost when making the data frame long")
}

length(unique(long_data$id))
nrow(data)
nrow(long_data)
covid_testing_data <- long_data

## fixing the dates of testing -- origin is again an artefact of xlsx
covid_testing_data %<>% dplyr::mutate(date_confirmed = as.Date(as.numeric(date_confirmed), 
                                                               origin = "1899-12-30"))


class(covid_testing_data$date_confirmed)

# check with Ilaria if this needs correcting anymore *** 
# fixing_dates <- function(date_vec) {
#   for(i in 1:length(date_vec)) {
#     if(date_vec[i] %in% seq.Date(from = as.Date("2020-03-03"), to = as.Date("2020-03-05"), 
#                                  length.out = 3)) {
#       date_vec[i] <- "2020-03-07"
#     } else {
#       date_vec[i] <- date_vec[i]
#     }
#   }
#   date_vec
# }
# fixing_dates(c(as.Date("2020-03-03"), as.Date("2020-03-04"), as.Date("2020-03-11"))) 
# fixing_dates("2020-03-02")
# # covid_testing_data$data_confirmed_real <- fixing_dates(covid_testing_data$data_confirmed)
#  
# covid_testing_data_fixing <- covid_testing_data %>%
#   dplyr::filter(date_confirmed > as.Date("2020-03-02")) %>%
#   dplyr::filter(date_confirmed < as.Date("2020-03-06")) %>%
#   dplyr::mutate(date_confirmed = as.Date("2020-03-07"))
# 
# covid_testing_others <- covid_testing_data %>%
#   dplyr::filter(date_confirmed < as.Date("2020-03-03") | date_confirmed > as.Date("2020-03-05")) 
# nrow(covid_testing_data_fixing) + nrow(covid_testing_others) == nrow(long_data)
# covid_data <- rbind(covid_testing_data_fixing, covid_testing_others) %>%
#   dplyr::arrange(id)

if(max(covid_testing_data$date_confirmed) != "2020-03-10") {
  stop("Error in formatting date -- max should be 10 March 2020")
}

## tests marked from 3 - 5 March were in fact not undertaken until at least the 6th -- correct to 7th
sort(unique(covid_testing_data$date_confirmed))

```

```{r only-cases}
## want to only filter through individuals who have at least one positive test
### from looking there are slight issues with household ids -- don't use these for now.
positive_ids <- (full_data[full_data$positive == TRUE,])$id

## covid data has some missing ids off positive people 
positives <- covid_testing_data %>% dplyr::filter(id %in% positive_ids)
positives <- positives[,c("id", "household_id", "date_confirmed", "test_result", "first_symptoms_date")]

```

```{r create-linelist}
## creformatting to get the kind of data frame we want - positive individuals and when they are first and last +ve/-ve 
covid_testing_data_pos <- positives %>% dplyr::group_by(id) %>%
  dplyr::filter(test_result == "Pos") %>%
  dplyr::mutate(first_positive = min(date_confirmed)) %>%
  dplyr::mutate(last_positive = max(date_confirmed))

covid_testing_data_neg <- positives %>% dplyr::group_by(id) %>%
  dplyr::filter(test_result == "Neg") %>%
  dplyr::mutate(first_negative = min(date_confirmed)) %>%
  dplyr::mutate(last_negative = max(date_confirmed))

covid_final <- dplyr::full_join(covid_testing_data_pos, covid_testing_data_neg, 
                                by = c("id","household_id")) %>%
  dplyr::left_join(covid_testing_data, by = c("id", "household_id"))

covid_final <- covid_final[,c("id", "household_id", "first_positive", "last_positive", 
                              "first_negative", "last_negative", "first_symptoms_date")] %>%
  dplyr::distinct() %>%
  dplyr::mutate(onset = first_symptoms_date) %>% 
  dplyr::mutate(onset = as.Date(as.numeric(onset), origin = "1899-12-30"))

## NAs introduced by coercion due to as.Date(as.numeric("NA"), origin = "1899-12-30") -- another excel consequence

```

```{r write-linelist}
linelist <- covid_final
## true means no duplicates
nrow(linelist) == length(unique(linelist$id))

```

```{r replace-contacts-with-anon-ids}

linelist_extras <- data.frame(id = missing_households$id, household_id = missing_households$household_id,
                              first_positive = rep(NA, length(missed_ids)), 
                              last_positive = rep(NA, length(missed_ids)), 
                              first_negative = rep(NA, length(missed_ids)),
                              last_negative = rep(NA, length(missed_ids)), 
                              first_symptoms_date = rep(NA, length(missed_ids)), 
                              onset= rep(NA, length(missed_ids)))

linelist %<>% as.data.frame()
linelist_extras %<>% as.data.frame()

names(linelist) == names(linelist_extras)

linelist <- rbind(linelist, linelist_extras)
linelist <- linelist[,c("id", "household_id", "first_positive", "last_positive", 
              "first_negative", "last_negative", "onset")]
 

write.table(linelist, "linelist.tsv", quote=FALSE, sep='\t', row.names = FALSE)

```

```{r write-rds-objects}
saveRDS(full_data, "full_data.rds")
saveRDS(linelist, "linelist.rds")
saveRDS(contact_final, "all_contact.rds")
saveRDS(direct_contact, "direct_contact.rds")


```